---
title: "CS2"
author: "David Nguyen, Austin Simeone, John Rodgers, Hannah Kosinovsky"
date: "December 4, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=TRUE}
library(reshape2)
library(ggplot2)
library(pander)
library(rcompanion)
library(GGally)
library(corrplot)
```

```{r dataload, echo=TRUE}
# load data from file in the data directory
data <- read.csv("./Data/API_17_DS2_en_csv_v2_10226244.csv", skip=4)
# generate rows for values for each year
data2 <- reshape(data, varying = 5:62, sep="", direction='long')
# rename columns
colnames(data2) <- c("Country","Code","Indicator","IndicatorCode","Value","Year","Id")
# isolate survival rate data into dataframe
survial.percent.male <- data2[data2$Indicator == "Survival to age 65, male (% of cohort)",]
# isolate literacy rate data into dataframe
literacy.youth.male <- data2[data2$Indicator == "Literacy rate, youth male (% of males ages 15-24)",]
# merge survival rate and literacy rate dataframes
survival.v.litrate.raw <- merge(survial.percent.male, literacy.youth.male, by=c("Code", "Year"))
# isolate specific variables from raw merged data
survival.v.literate <- survival.v.litrate.raw[c(3,1,2,6,11)]
# rename columns
colnames(survival.v.literate) <- c("Country","Code","Year","survival.percent","literate.rate")
# exclude na values
survival.v.literate <- na.omit(survival.v.literate)
# generate log of survival percent
survival.v.literate$log.survival.percent <- log(survival.v.literate$survival.percent)
# generate log of literate rate
survival.v.literate$log.literate.rate <- log(survival.v.literate$literate.rate)
``` 

The variables used for linear regression at literacy rate in youth males (age 15-24) and the survival rate of males to age 65. A summary of those variables and their distribution among the sample are below.
```{r variable-summaries}
pander(summary(literacy.youth.male$Value))
plotNormalHistogram(literacy.youth.male$Value)
pander(summary(survial.percent.male$Value))
plotNormalHistogram(survial.percent.male$Value)
```

A plot of the log of the survival rate and the log of the literacy rate shows that the two variables have a linear correlation.
```{r plot, echo=TRUE}
# plot surivival rate versus literate rate
ggplot(survival.v.literate, aes(x=log.literate.rate, y=log.survival.percent)) + geom_point(shape=1) + geom_smooth(method=lm)
```


```{r residuals, echo=TRUE}
# generate linear model of survival rate vs literate rate
s.v.l.lm <- lm(log.survival.percent ~ log.literate.rate, data=survival.v.literate)
# generate residuals
s.v.l.m.res <- resid(s.v.l.lm)
# plot residuals
plot(survival.v.literate$log.literate.rate, s.v.l.m.res, ylab="Residuals",xlab="log.literate.rate", main="Residuals vs log of literate rate")
```

```{r studentresiduals, echo=TRUE}
s.v.l.m.stud <- rstudent(s.v.l.lm)
hist(s.v.l.m.stud, freq=FALSE, main="Distribution of Studentized Residuals", xlab="Studentized Residuals")
xfit <- seq(min(s.v.l.m.stud)-1,max(s.v.l.m.stud)+1,length=40)
yfit <- dnorm(xfit)
lines(xfit,yfit)
```
```{r qqplot, echo=TRUE}
par(mfrow=c(1,2))
qqnorm(s.v.l.m.stud)
qqline(s.v.l.m.stud)
```
```{r summary, echo=TRUE}
pander(summary(s.v.l.lm))
```
$\hat{\mu}(log(Survival Percent)|log(Literate Rate)) = 0.126 + 0.894 * log(Literate Rate)$

The doubling of the literate rate if male youth results in a change of 85% increase in the rate of males surviving to 65. 

About 49.5% of the variation in the log of Survival Rate is explained by the log of the Literate Rate.


```{r summary_of_life_expectancy}
mothers.postion.maternity <- data2[data2$IndicatorCode == 'SG.MMR.LEVE.EP',]

female.life.expectancy<- data2[data2$IndicatorCode == "SP.DYN.LE00.FE.IN",]

male.life.expectancy<- data2[data2$IndicatorCode =="SP.DYN.LE00.MA.IN",]

female.participation <- data2[data2$IndicatorCode =="IC.FRM.FEMO.ZS",]

female.manager <- data2[data2$IndicatorCode =="IC.FRM.FEMM.ZS",]

data3 <- Reduce(function(x,y) merge(x = x, y = y, by = c("Code", "Year")), 
       list(mothers.postion.maternity, 
            female.life.expectancy,
            male.life.expectancy,
            female.participation,
            female.manager))

data3 <- data3[complete.cases(data3),]

names(data3)[c(6,11,16,21,26)] <- c('a','b','c','d','e')

#MULTICOLINEARITY
explanatory_variables <- data3[c(6,11,16,21,26)]

#visually 
ggpairs(explanatory_variables)

#complete data only
#correlation of each feature and another
m <- cor(explanatory_variables)

#plots correlation matrix
corrplot(m, method = c('number'))

fit <- lm(b~a+d+e, data = data3)

summary(fit)

summary(data3[c(6,11,16,21,26)])

lapply(data3[c(6,11,16,21,26)], plotNormalHistogram)

```






